\documentclass[sigconf,review,screen]{acmart}




\usepackage{wasysym}
\usepackage{graphicx}
\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{etex}

\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{listings}
\usepackage[english]{babel}
\usepackage{multirow}
\usepackage{times}
\usepackage{tabularx}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[export]{adjustbox}
\usepackage[utf8]{inputenc}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}%
\usepackage{etoolbox}
\usepackage{tcolorbox}
\usepackage{xcolor}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{bbm}
\usepackage[flushleft]{threeparttable}
\DeclareMathOperator{\poly}{\textup{poly}}

\usepackage{etoolbox}
\usepackage{marginnote}
\usepackage{soulutf8}
\soulregister{\cite}{7} % Register \cite command with soul
\soulregister{\emph}{7} % Register \cite command with
\soulregister{\autoref}{7} % Register \cite command


\makeatletter
\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
\patchcmd{\SOUL@ulunderline}{\dimen@}{\SOUL@dimen}{}{}
\patchcmd{\SOUL@ulunderline}{\dimen@}{\SOUL@dimen}{}{}
\patchcmd{\SOUL@ulunderline}{\dimen@}{\SOUL@dimen}{}{}
\patchcmd{\SOUL@ulunderline}{\dimen@}{\SOUL@dimen}{}{}
\newdimen\SOUL@dimen
\makeatother
%
\sethlcolor{yellow!20}

% \DeclareRobustCommand{\update}[3][0em]{\marginnote{#2}[#1]\hl{#3}}
\DeclareRobustCommand{\update}[3][0em]
\makeatother


\newcommand{\code}[1]{{\textsf{{\small #1}}}}
% \newcommand{\todo}[1]{\textcolor{red}{\textit{#1}}}
\newcommand{\todomargin}[1]{\marginpar{\textcolor{red}{\textit{#1}}}}

\definecolor{forestgreen}{HTML}{228C22}

\renewcommand{\labelitemi}{$\circ$}
\renewcommand{\labelitemii}{$-$}

\def\Snospace~{\S{}}
\renewcommand*\sectionautorefname{\Snospace}
\renewcommand*\subsectionautorefname{\Snospace}
\renewcommand*\algorithmautorefname{Alg.}
\newcommand*\exampleautorefname{Example}
\newcommand*\lemmaautorefname{Lemma}
%\newcommand*\theoremautorefname{Theorem}
\newcommand*\propertyautorefname{Property}
\newcommand*\problemautorefname{Problem}
\newcommand*\definitionautorefname{Def.}
\renewcommand*{\algorithmcflinename}{line}
\renewcommand\equationautorefname{Eq.}
\renewcommand*\figureautorefname{Fig.}


\SetAlgoSkip{}

\newcommand{\exampleboxh}[2]{
\begin{tcolorbox}[colback=white,
	float,
	floatplacement=h,
	enlarge top by=-.6em,
	enlarge bottom by=-.6em,
	colframe=gray!75!black,
	title=\scshape\fontseries{sb}\selectfont#1,
	boxsep=3pt,
	left=1pt,
	right=1pt,
	top=0pt,
	bottom=0pt,
	toptitle=0pt,
	bottomtitle=0pt,
	lefttitle=1pt,
	righttitle=1pt,
	]{\small%
#2%
}
\end{tcolorbox}%
}

\newcommand{\examplebox}[2]{
\begin{tcolorbox}[colback=white,
	float,
	floatplacement=t,
	enlarge bottom by=-.6em,
	colframe=gray!75!black,
	title=\scshape\fontseries{sb}\selectfont#1,
	boxsep=3pt,
	left=1pt,
	right=1pt,
	top=1pt,
	bottom=1pt,
	toptitle=0pt,
	bottomtitle=0pt,
	lefttitle=1pt,
	righttitle=1pt,
	]{\small%
		#2%
	}
\end{tcolorbox}%
}

\newtheorem{corollary}{Corollary}
\newtheorem{problem}{Problem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{property}{Property}
\newtheorem{theorem}{Theorem}

\lstset{ %
	belowskip=-2em,
	backgroundcolor=\color{white},   % choose the background color; you must
	basicstyle=\footnotesize\ttfamily,        % the size of the fonts that are
	%used
	%for the code
	breakatwhitespace=false,         % sets if automatic breaks should only
	%happen at whitespace
	breaklines=true,                 % sets automatic line breaking
	captionpos=b,                    % sets the caption-position to bottom
	commentstyle=\color{Brown},    % comment style
	deletekeywords={...},            % if you want to delete keywords from the
	%given language
	escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
	extendedchars=true,              % lets you use non-ASCII characters; for
	%8-bits encodings only, does not work with UTF-8
	frame=,                    % adds a frame around the code
	keepspaces=true,                 % keeps spaces in text, useful for keeping
	%indentation of code (possibly needs columns=flexible)
	keywordstyle=\color{ACMDarkBlue},       % keyword style
	language=,                 % the language of the code
	morekeywords={*,RETURN,SEQ,WITHIN,PATTERN,NOT,
		WHERE,AND,NSEQ,ms,h, min, range},
	%
	%if you want to add more keywords to the set
	numbers=none,                    % where to put the line-numbers; possible
	%values are (none, left, right)
	numbersep=5pt,                   % how far the line-numbers are from the
	%code
	numberstyle=\tiny\color{gray}, % the style that is used for the line-numbers
	rulecolor=\color{black},         % if not set, the frame-color may be
	%changed
	%on line-breaks within not-black text (e.g. comments (green here))
	showspaces=false,                % show spaces everywhere adding particular
	%underscores; it overrides 'showstringspaces'
	showstringspaces=false,          % underline spaces within strings only
	showtabs=false,                  % show tabs within strings adding
	%particular
	%underscores
	stepnumber=1,                    % the step between two line-numbers. If
	%it's
	%1, each line will be numbered
	stringstyle=\color{RedOrange},     % string literal style
	tabsize=2,                       % sets default tabsize to 2 spaces
	title=\lstname,                   % show the filename of files included
	%with
	%\lstinputlisting; also try caption instead of title
}



\hypersetup{
	pdftoolbar=true,        % show Acrobat?s toolbar?
	pdfmenubar=true,        % show Acrobat?s menu?
	pdffitwindow=false,     % window fit to page when opened
	pdfstartview={FitH},    % fits the width of the page to the window
	pdftitle={},    % title
	pdfauthor={},     % author
	pdfsubject={},   % subject of the document
	pdfcreator={},   % creator of the document
	pdfproducer={}, % producer of the document
	pdfkeywords={}, % list of
	%keywords
	pdfnewwindow=true,      % links in new window
	colorlinks=true,       % false: boxed links; true: colored links
	  linkcolor=Brown,          % color of internal links (change box color
%	with
%	linkbordercolor)
	  citecolor=OliveGreen,        % color of links to bibliography
	  filecolor=magenta,      % color of file links
	  urlcolor=NavyBlue           % color of external links
}

\newcommand{\sys}{DISCES\xspace}

\renewcommand{\UrlFont}{\ttfamily\scriptsize} %?
\newcommand{\sstitle}[1]{\smallskip\noindent\textbf{#1.\/}}
\newcommand{\sstitlenoskip}[1]{\noindent\textbf{#1.\/}}

\newcommand{\change}[1]{\textcolor{blue}{#1}}


\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\argmin}{argmin}

\newcommand\Mark[1]{\textsuperscript#1}
\newcommand{\True}{\mathit{True}}
\newcommand{\False}{\mathit{False}}
\def\ua{\Mark{*}}
\def\ub{\Mark{\dag}}
\def\uc{\Mark{\#}}
\def\ud{\Mark{+}}

\begin{document}
\title{\sys{}: Systematic Discovery of Event Stream Queries}
\subtitle{Technical Report\vspace{1cm}}


%% The "author" command and its associated commands are used to define the
%%authors and their affiliations.
\author{Rebecca Sattler}
\affiliation{%
 \institution{Humboldt-Universit\"at zu Berlin}
 \city{Berlin}
 \country{Germany}
}
\email{rebecca.sattler@informatik.hu-berlin.de}

\author{Sarah Kleest-Meißner}
\affiliation{%
 \institution{Humboldt-Universit\"at zu Berlin}
 \city{Berlin}
 \country{Germany}
}
\email{kleemeis@informatik.hu-berlin.de}

\author{Steven Lange}
\affiliation{%
 \institution{Humboldt-Universit\"at zu Berlin}
 \city{Berlin}
 \country{Germany}
}
\email{langestx@hu-berlin.de}

\author{Markus L. Schmid}
\affiliation{%
	\institution{Humboldt-Universit\"at zu Berlin}
	\city{Berlin}
	\country{Germany}
}
\email{MLSchmid@MLSchmid.de}

\author{Nicole Schweikardt}
\affiliation{%
 \institution{Humboldt-Universit\"at zu Berlin}
 \city{Berlin}
 \country{Germany}
}
\email{schweikn@informatik.hu-berlin.de}

\author{Matthias Weidlich}
\affiliation{%
 \institution{Humboldt-Universit\"at zu Berlin}
 \city{Berlin}
 \country{Germany}
}
\email{matthias.weidlich@hu-berlin.de}



\begin{abstract}
The continuous evaluation of queries over an event stream provides the
foundation for reactive applications in various domains.
Yet, \update{M1\\R1O1}{knowledge of queries that
detect
distinguished event
patterns that are potential causes of the situation of
interest is often not directly available.
However, given a
database of finite, historic
(sub-)streams that have been gathered whenever a situation of interest was
observed, one may aim at automatic discovery of
the respective queries.} Existing algorithms for event query discovery
incorporate ad-hoc design choices, though,  and it is unclear how their
suitability for a database shall be assessed.

In this paper, we address this gap with \sys{}, an algorithmic framework for
event query discovery. \sys{} outlines a design space for discovery algorithms,
thereby making the design choices explicit. We instantiate the
framework to derive four specific algorithms, which all yield correct and
complete results, but differ in their runtime sensitivity. We therefore also
provide guidance on how to select one of the algorithms for a given database
based on a few of its essential properties. Our experiments using simulated and
real-world data illustrate that our algorithms are indeed tailored to databases
showing certain properties and solve the query discovery problem several
orders of magnitude faster than existing approaches.
\end{abstract}

\maketitle


\input{introduction.tex}
\input{problem_statement.tex}
\input{framework.tex}
\input{discovery_algorithms.tex}

\input{instantiation.tex}
\input{experiments.tex}
\input{related_work.tex}
\input{conclusions.tex}




\begin{acks}
 This work was funded by the German Research
Foundation (DFG), CRC 1404: "FONDA: Foundations of
Workflows for Large-Scale Scientific Data Analysis".

\end{acks}

\clearpage

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\clearpage
\input{appendix.tex}

\end{document}
\endinput
